---
title: "Assignment 2 Task 2 - Parameter Estimation with Purr"
author: "Logan Ossentjuk"
date: "2/5/2022"
output: 
  html_document:
    code_folding: hide
---
# Overview 



```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE,  message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(broom)
library(kableExtra)
library(cowplot)
library(modelr)
library(purrr)
```

```{r}
lizards <- read_csv(here("data", "lizard.csv")) %>% 
  mutate(sex = case_when(sex == 'M' ~ 'Male',
                         sex == 'F' ~ 'Female')) %>% 
  clean_names() %>% 
  #Log Transformation
  mutate(log_weight = log(weight)) %>% 
  mutate(log_length = log(sv_length)) %>% 
  drop_na()
```

\[
W=a(SVL)^b
\]

*Weight is given by W, snout to vent length by SVL, and a and b are the parameters that need to be fitted.*

```{r, include=FALSE}

#Create Model 

lizards_model <- function(a, SVL, b) {
  W = a*(SVL)^b
  return(W)
}

# Generate parameter guesses from coefficients 

model_guess <- lm(log_weight ~ log_length, data = lizards)
```


```{r}
# Supply NLS with regression coefficients

cf <- coefficients(model_guess)

#Parameter b (length) guessed as 2.54, a calculated using `exp()`
lizards_nls <- nls(weight ~ lizards_model(a, sv_length, b),
                  data = lizards,
                  start = list(a = exp(cf[1]),
                               b = cf[2]))
```


```{r}
### Create Table 
tidy_nls <- tidy(lizards_nls)

tidy_nls$p.value <- ifelse(tidy_nls$p.value < .001, paste("< .001"))

tidy_nls %>% 
  select(-statistic) %>% 
kable(col.names = c("Terms", "Coefficients", "Standard Error", "p Value"), caption = "**Table 1**", digits = 3) %>% 
  kable_classic(bootstrap_options = "striped", full_width = FALSE) 
```

```{r}
model_predict <- augment(lizards_nls)

# Plotting predicted data given NLS
ggplot() +
   geom_point(data = lizards, 
             aes(x = sv_length, 
                y = weight, 
                color = sex)) +
   geom_line(data = model_predict, 
             aes(x = sv_length,
                 y = .fitted,
                 color = "Model Predictions")) +
  scale_color_manual(values = c("pink", "skyblue", "black")) +
  theme_classic() +
  theme(legend.title = element_blank()) +
  labs( x = "SV Length (mm)", y = "Weight (g)", 
        caption = "**Figure 1.** Displays the predicted non-linear square model for SV Length (mm) versus Weight (g). Female lizards are represented by pink points, Male by blue and the black line represents the 
trained NLS model to predict Weight given SV Length.")
```

**Figure 1.** Displays the predicted non-linear square model for SV Length (mm) versus Weight (g).
Female lizards are represented by pink points, Male by blue and the black line represents the 
trained NLS model to predict Weight given SV Length.



```{r}
### Preparation for Table 

### Filter for male Western Whiptails 

lizards_m_whip <- lizards %>% 
  filter(sex == "Male", spp == "CNTI")

#Calculate RMSE 
lizards_rmse <- rmse(lizards_nls, data = lizards)
lizards_m_whip_rmse <- rmse(lizards_nls, data = lizards_m_whip)

### Create model for subset
guess_m_whip <- lm(log_weight ~ log_length, data = lizards_m_whip)

#tidy_guess_m_whip <- tidy(guess_m_whip)

m_whip_cf <- coefficients(guess_m_whip)

m_whip_nls <- nls(weight ~ lizards_model(a, sv_length, b),
                  data = lizards_m_whip,
                  start = list(a = exp(m_whip_cf[1]),
                               b = m_whip_cf[2]))

tidy_m_whip_nls <- tidy(m_whip_nls) # tidy output of model
```


```{r}
### Table of Male Whiptail NLS Model
tidy_m_whip_nls %>%  
  mutate(p.value = case_when(
    p.value < 0.001 ~ "<0.001",
    p.value > 0.001 ~ (as.character(round(p.value, 3))))) %>% 
  kable(col.names = c("Term", 
                      "Estimate",
                      "Standard Error", 
                      "Statistic", 
                      "p Value"), caption = "**Table 2**", digits = 3) %>% 
  kable_classic(bootstrap_options = "striped", full_width = FALSE)

```


```{r}
#Generate predictions based on model

m_whip_predict <- augment(m_whip_nls)

#Create a complete data frame
m_whip_complete <- data.frame(lizards, model_predict) %>% 
  filter(sex == "Male", spp == "CNTI")

#Calculate RMSE
m_whip_rmse <- rmse(m_whip_nls, data = lizards_m_whip)
lizards_complete_rmse <- rmse(lizards_nls, data = lizards_m_whip)

# Graphing model fits given NLS
ggplot() +
  geom_point(data = lizards_m_whip, 
             aes(
               x = sv_length,
               y = weight,
               color = "Actual Data")) +
  geom_line(data = m_whip_predict, 
            aes(
            x = sv_length,
            y = .fitted,
            color = "Male Whiptails Model")) +
  geom_line(data = m_whip_complete, 
            aes(
            x = sv_length,
            y = .fitted,
            color = "All Lizards Model")) +
  scale_color_viridis_d() +
  theme_classic() +
  labs( x = "SV Length (mm)", y = "Weight (g)") +
  theme(legend.title = element_blank())
```

**Figure 2:** Compares true lizards weights (red) in comparison to SV Length. The weights predicted by the *Male Whiptail Model* (yellow ) and the *All Lizards Model* (blue line) =  previously shown in Figure 1 represent the populations in which these models were trained upon. Statistical analysis reveals that the *Male-Whiptails Model* better predicts actual weights (**RMSE** =`r round(m_whip_rmse, 2)`) better than the *All Lizards Model*, (**RMSE** = `r round(lizards_complete_rmse, 2)`). Given the lower RMSE of the *Male-Whiptails Model*, when predicting Weight as a function of SV Length, models based upon the Male Western Whiptail population are preferred. 


**Data Citation** Lightfoot, D. and W.G. Whitford. 2020. Lizard pitfall trap data from 11 NPP study locations at the Jornada Basin LTER site, 1989-2006 ver 37. Environmental Data Initiative. https://doi.org/10.6073/pasta/4a6e258fb49c31e222ecbbcfd128967f
